Regional and Urban Economics

### 2023-02-19 Retail-Agglomeration_abstract_introduction_theori_econometric_approach.tex
...

\begin{document}
\maketitle
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, borderline west={3pt}{0pt}{shadecolor}, frame hidden, interior hidden, boxrule=0pt, enhanced, breakable]}{\end{tcolorbox}}\fi

@@

\texttt{The\ topic\ of\ research\ (What\ was\ studied)}

\texttt{The\ purpose\ (What\ was\ done)}

\texttt{The\ methodology\ (how\ the\ research\ was\ done)}

\texttt{Main\ findings\ (what\ was\ found)}

\texttt{Policy\ and\ implications\ (significance\ of\ findings)}

@@

\texttt{A\ brief\ description\ of\ the\ problem\ and\ its\ importance}

\texttt{Statement\ of\ purpose\ or\ what\ was\ done}

\texttt{A\ brief\ description\ of\ main\ results}

\texttt{Description\ of\ methodology,\ main\ challenges\ and\ solutions}

\texttt{Discretion\ of\ the\ papers\ contribution\ in\ relation\ to\ previous\ research}

\texttt{Organization\ of\ paper}

@@

Location theory gives regional economics its scientific disciplinary
identity and constitutes its theoretical methodological core
(\textbf{capello2011?}). It has typically microeconomic foundation and
uses theoretical models as well as adopting a statistical and
geographical approach (\textbf{capello2011?}). Furthermore, the theory
uses the concept of externalities in the spatial distribution of
activities, thereby laying the territorial bases for dynamic approach to
economic growth (\textbf{capello2011?}).

Regional growth theory involves spatial aspects of economic growth and
territorial distribution of income (\textbf{capello2011?}). It also
involves generating geographical advantages, in terms of easy or
difficult access to a particular area (\textbf{capello2011?}).

@@

Harald Hotelling's locational equilibrium is determined by a logic of
profit maximization whereby each producer controls its own market area.
Productivity advantages of cities and urban clusters with a high density
of firms increase profit by attracting a larger number of potential
customers, and more productive workers (\textbf{capello2011?}).
Furthermore, the attractiveness of a central location increase the cost
of rent.

Alonso's bid rent model indicates the most profitable location for
firms. Closer to the centre with agglomeration attributes or in rural
areas with spacial monopoly and low rent. In gravitational models, the
attractiveness of the retail location, represent the size of the retail
centre. Furthermore, it depends on the variety of goods which can be
purchased at the same location (\textbf{mccann2013?}).

The model of potential has the capacity to measure the potential of
attractiveness to a place. Bigger cities or more heavily populated areas
have a stronger gravitational force. A possible indicator to predict
places of growth potential for dentist practice or shopping malls
(\textbf{capello2015?}).

@@

The model assume that given the location of producers, and given demand
uniformly distributed geographically (in linear or circular form) the
market is divided into areas within each of which there operates a
single firm in a duopoly environment. Furthermore, no relocation costs
and demand only depend on location choice (\textbf{capello2015?}). The
location game starts off with the total market of AB, firm A in the
middle of location A and firm B starts in the middle of location B. One
firm starts relocating closer to the other to take some of the customers
in the other market area (\textbf{capello2015?}). The other respond by
doing the same and the game continues until both end up in the middle of
the market on the broader of AB. The end of the game is the position
where neither can increase sales volume by moving position
(\textbf{capello2015?}).

@@

Marshall (1920) broadly divides externalities within agglomeration in
three main categories potentially drive sales. Firstly, knowledge
spill-over within industries or product specific technological
knowledge. Furthermore, market transactions in terms of value chain
transactions with industry-specialized buyers and suppliers. Lastly,
competition for specialized production factors such as labour and
product market competition (\textbf{nielsen2021?}).

There are solidly established conclusions regarding the existence of
agglomeration economies (\textbf{puga2010?}). However, less proof of
their estimated magnitude. Hence, identifying the causes of
agglomeration economies, is proving more difficult (\textbf{puga2010?}).
Nevertheless, there is a large theoretical literature that develops
these mechanisms (\textbf{puga2010?}). (\textbf{duranton2004?}) discuss
these classifications and identify \emph{learning, sharing} and
\emph{matching} as the main causes of agglomeration economies.

A larger market allows for a more efficient sharing of local
infrastructure and facilities. Therefore, a variety of intermediate
input suppliers, or a pool of workers with similar skills
(\textbf{puga2010?}). Despite higher rent the dental industry and
shopping malls, seams to reap higher benefits in more populated areas as
they are dependent on being located where there is a higher volume of
patients in order to drive sales. The attraction for the consumers and
users of public facilities is overall cost reduction
(\textbf{puga2010?}). Hence, the larger the population sharing
facilities the lower the cost per user (\textbf{puga2010?}). Presumably,
industrial factories and business clusters are more dependent on being
close to raw materials and industrial action.

Furthermore, a larger market also allows for a better matching between
employers and employees, improved chances of finding suitable and better
quality of matches (\textbf{puga2010?}). Shopping malls require skilled
workers to drive sales. However, they are not so dependent on highly
educated workers as dentists whom according to resent study, tend to
prefer bigger cities (as they are highly educated)
(\textbf{davis2020?}). More-so, cites provide a constant market for
specialized skills and more productive work force (\textbf{puga2010P?})
Perhaps, a possibility for higher wages being compensated by more
efficient workers in bigger agglomeration economies. That said the
Norwegian cities might differ from the american cites in the study, as
the population and clusters are no where near the same size. Lastly, a
larger market can also facilitate learning by promoting the development
and widespread adaptation of new technologies and business practice
(\textbf{puga2010?}).

Interactions with experienced workers helps acquire valuable skills.
Experienced workers remain in cities to share the rent of this learning
process (\textbf{puga2010?}). Besides this purposeful transition of
knowledge, the information literature on learning in cities has also
emphasized the unintended casual flow of information facilitated by big
cities.'' (\textbf{puga2010?})

@@

The Alsonso model demonstrate geographical locations tied up to
location. Furthermore, It is an urbanized formulation of the nineteenth
century von Thünen model (\textbf{capello2015?}). In this case used to
get an indicating of where the most profitable location for dental
practise.

Alonso assumes the existence of a city that cannot be build
instantaneously, and therefore of an effective rent-curve from the city
centre to the periphery. Furthermore, it determines the location for a
new firm willing to locate in the city and the profit the firm can
obtain. In some cases different from the normal or average price
competition (\textbf{capello2015?}).

As the Von Thünen assume a uniformed space where all land is equally
fertile, this model envisages a city, endowed with infrastructure which
cover the entire city in all directions whereas unit transport costs are
constant in space. More-so, the town or city has a single centre point.
The city centre of business district (CBD) is generally defined as the
most attractive for all firms. (\textbf{capello2015?}). The model also
assume perfect competition and unlimited demand, in other words it is
supply-orientated (\textbf{capello2015?}). Furthermore, demonstrating a
specific production function with fixed coefficients and constant
returns to scale.

Rent obtained in the model is the remainder left after transport,
production costs and profit has been subtracted from revenues
(\textbf{capello2015?}). The bid-rent curve will in this case
demonstrate the prise of rent within Bank and Financial locations,
Retail locations, and industrial locations.

@@

The illustration shows three categories: The most expensive rent closest
to the centre is the Bank and finance district, followed by the retail
agglomeration areas and industrial sites furthest away from the central
point. Dental practices are service orientated thus depending on
population density and central activities. Opening hours are relay
outside of office hours. Production cost does not appear to be important
in this case since dentist is a public service. Hence, not dependent of
regular transportation of sales goods. The rent realized by the
landowner is the envelope of the three rent supply curves
(\textbf{capello2015?}).

From a Marshallian way of thinking dentists should cluster somewhere in
this pink area to reap maximum sales volume. Furthermore, they may also
increase their profits further out in to other retail agglomeration
area. The attractiveness might be high in these places due to economies
of scope. As the positive benefits from the Marshallian agglomeration
attributes play out. Subsequently, increasing profit by trading higher
transportation cost for lower rent. In addition fewer dentists to
compete with supporting the hotellings game theory whereas companies
more to achieve spatial monopoly (\textbf{mccann2013McCann?}).

@@

The centre attracts firms, people and activities. Furthermore, it
influence the central points in diverse ways such as commuter movements,
diffusion of knowledge, corporations and network and personal
relationship (\textbf{capello2015?}). More-so, they are generated by the
intensity of the flows generated between places (\textbf{capello2015?}).
The gravitational pull to a place of interest for a dentist will
determine how much sales can be generated by locating there. Several
locations can then be measures against each other to determine profit
maximisation.

These models has the potential to measures the future attractiveness of
a place. Changes in infrastructure such as easier access to a place
might change the attractiveness of a place or move the flow of people.
Subserviently, increasing the gravitationally pull between the two
places. Furthermore, The model uses population and distance to measure
the attractiveness and relations between two places
(\textbf{capello2015?}). Furthermore, these models have predictive
capacity and can be used to estimate the potential impact. In
particular, location of new productive activity in an area
(\textbf{capello2015?}). For instance predict where to build a shopping
centre (\textbf{capello2015?}). By taking these factors in to account
dentist could predict the most profitable locations in years to come by
predicting the amount of people that will move from one city to another
or tow different points of interest within a city.

@@

Town and cities have long had shopping districts in which stores are
concentrated. To some extent such centre points were the result of
uncoordinated store-location decisions (\textbf{brueckner2011?}).
Nevertheless, not coincidentally according to Hotelling's theory (1929).
He describes the concentration of retail stores by a market desire to
create a Nash equilibrium. At this point all retailers expected sales
maximisation outcome cannot be improved by changing position.
Contradictory to the assumption that consumers will shop at the nearest
retailer, they are willing to travel significantly further to reach
certain central places. Marshall (1920) points out another explanation
for this in his seminal work on agglomeration. That being,intra-industry
knowledge spillovers as mechanisms leading to formation of industry
cluster (\textbf{nielsen2021?}). In this case the dental industry
information spillovers are bounded in geographical space. Hence, the
need to locate in close proximity. That being the latest equipment, in
order to work faster or the right personnel to attract more patients.
Another way to drive sales in this area could be by dentist
specialisation and patients referrals between dentists.

Another important externalities expressed in Marshall (1920) theory is
market transactions. Whereas, geographical proximity reduces transaction
cost through several mechanism (\textbf{nielsen2021?}). The market
transactions can be extended to the endogenous location of specialized
human capital (\textbf{nielsen2021?}). ``It is well known among
economists and other social scientists that large cities have
disproportionately large shares of highly educated workers, and the
trend has been growing in recent decades.'' (\textbf{bakerBaker?}).
Presumably, larger cities would then be more attractive to dentists in
terms of cost and quality of specialised labour in smaller towns.
Especially, those trying to achieve spatial monopoly end up loosing
profit by overpaying for labour or in worst case scenario not finding
enough dentists labour and end up compromising on quality. In both cases
growth potential would be compromised.

Marshall also points out the competition within the geographical
location. Hence, the marketplace and business districts give consumers
low search and switching costs (\textbf{capello2015?}). The fact that
owners of stores seem to prefer highly concentrated shopping. More-so,
it suggests that the sales volumes outweigh the loss attributed to
greater price competition@capello2015.

@@

Owners of shopping malls ``orchestrate'' the process of retail
agglomeration that happens naturally in towns and cities
(\textbf{brueckner2011?}). Naturally, the same retail agglomeration and
Nash equilibrium may occur in shopping malls if not more intensely as
the shops are in closer proximity. Furthermore, the potential for more
knowledge spillovers, higher volume of market transaction and more
competition. When retail agglomeration is orchestrated by the owner of a
shopping mall, the strength and direction of such externalities are
taken into account (\textbf{brueckner2011?}). Much so, because the price
of rent is dependent on the overall revenue of a shopping mall. Hence,
the owner of the mall wants to choose the mix of stores and their sizes
taking these externalities, attributes and anchor stores attractiveness
into account (\textbf{brueckner2011?}).

@@

Spatial data analysis is a rapidly growing area in Statistics
(\textbf{haslett1992?}). The Journal of the statistics society define
such data as being objects in space, which objects have physical
location. Furthermore, the analyses are spatial if these locations are
relevant to the interpretation of the data (\textbf{haining1998?}).

The examination of the data involves the examination of data collected
from q-gis maps;
\texttt{it\ will\ involve\ little\ more.\ Graphical\ techniques\ for\ examining\ such\ maps\ are\ thus\ a\ central\ part\ of\ the\ methodology.\ A\ number\ of\ such\ techniques\ are\ discussed."}
(\textbf{haslett1992?}).

@@

The bid rent models are under the assumption that Dentists use spatial
completion to maximise profit. Nevertheless, the cost of dental care in
Norway can vary depending on a number of other factors such as the type
of treatment, the location of the dental practice and the dentist
experience and qualifications. The Norwegian Dental Association rate the
average cost of check ups is between four to eight hundred Norwegian
Krone's {``Tannlegeforeningen''} (n.d.). Bertrand model also take price
as competitive advantage in to consideration.

@@

Shopping malls have a relatively short history in Norway, with the first
shopping mall opening in the late sixties. Prior to the most retail
activities took place in the city centre. Furthermore, these malls
typically featured a mix of shops, restaurants and entertainment
options, and they were often located in suburban areas or on the
outskirts of cities {``Home - NCSC''} (n.d.). In resent years, there has
been a trend towards larger and more comprehensive shopping malls in
Norway {``Home - NCSC''} (n.d.).

However, the growth of shopping malls in Norway has not been without
controversy. Some critics argue that the development of large shopping
malls has led to the decline of traditional shopping streets and local
businesses, while other raise concerns about the environmental impact of
these large scale commercial developments {``Kan Vin Og Sprit Skape
Byutvikling?''} (2014).

Information about Wine monopoly locations has also been studied in this
paper. As the wine monopoly are anchor stores they may affect the
attractiveness of the shopping mall {``Kan Vin Og Sprit Skape
Byutvikling?''} (2014). The Norwegian government regulates the
placements of wine monopoly stores in shopping malls to prevent
completion with town centres omsorgsdepartementet (2022). More
importantly, the government may take in to account the potential impact
of exciting retail centres in the region. This regional economic
approach aim to promote sustainable development and balance economic
growth across the country {``Kan Vin Og Sprit Skape Byutvikling?''}
(2014).

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-home-n}{}}%
{``Home - NCSC.''} n.d.
\href{https://ncscnordic.org/,\%20https://ncscnordic.org/}{https://ncscnordic.org/,
https://ncscnordic.org/}.

\leavevmode\vadjust pre{\hypertarget{ref-kanvin2014}{}}%
{``Kan Vin Og Sprit Skape Byutvikling?''} 2014.
\url{https://www.aftenposten.no/kultur/i/oEy0/kan-vin-og-sprit-skape-byutvikling}.

\leavevmode\vadjust pre{\hypertarget{ref-omsorgsdepartementet2022}{}}%
omsorgsdepartementet, Helse-og. 2022. {``Vinmonopolordningen.''}
\emph{Regjeringen.no}. regjeringen.no.
\url{https://www.regjeringen.no/no/tema/helse-og-omsorg/psykisk-helse/innsikt/vinmonopolordningen/id2528183/}.

\leavevmode\vadjust pre{\hypertarget{ref-tannlege}{}}%
{``Tannlegeforeningen.''} n.d. \url{https://www.tannlegeforeningen.no/}.

>>>>

\begin{document}
\maketitle
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, enhanced, boxrule=0pt, frame hidden, interior hidden, breakable, sharp corners]}{\end{tcolorbox}}\fi

@@

There are several factors to take in to consideration when deciding
where a dental practice should be located. Spatial competition is a
critical consideration in regional economics. Some of the factors
influencing the decision making is retail agglomeration attributes,
spatial monopoly and the attractiveness of a particular place of
interest. This study aims to examine the dental industry and find the
strategy most profitable for the dental industry.

The study utilize a quantitative research approach, employing spacial
explanatory data analysis techniques to determine the location decisions
of firms. Geographical data was collected from locations in Norway
showing all dentists, central places, retail agglomeration and other
points of interest such as the government regulated wine monopoly.
Combined with regional income data a spatial regression analyses was
used to identify the factors influencing the location decisions of
firms.

Furthermore, the result indicates the distance to retail agglomeration
areas and central places playes a significant role in the location
decisions for firms. Firms are more likely to locate closer to these
areas, which offer access to a larger customer base and lower cost of
transportation. However, the results also suggest other factors, such as
market size and completion also influence the location decisions for
firms.

The studies findings have particular implications for policy-makers,
developer, and dental practises seeking to maximize profits through
strategic decisions. By understanding the importance of proximity to
retail agglomeration areas and central places, these stakeholders can
make informed decisions that benefit the local economy and improve
competitiveness of the region. Perhaps more importantly prevent
agrassive retail agglomeration areas from ``killing down town''.

@@

The theoretical framework for this study draws on Marshall's
agglomeration principles, which emphasize spatial aspects of economic
growth. Hotelling's game theory is used to explain spatial competition
by location. Furthermore, adding the hybrid model whereas firms compete
both in location and price, with consumers having a transportation cost
and firms producing differentiated products. Alsonso's bid-rent model
and gravitational models also provides insights into how dentists or
other retailer makes location decisions in the context of regional
growth.

The dental industry provides an interesting industry to study for this
research because of its unique characteristics. Dental practices require
a high degree of specialization and expertise. At the same time, they
need to be easily accessible to consumers. Shopping malls and central
places offer a continent location for dental practices that said
competition for prime locations can be intense and push some in to
spatial monopoly.

The result of this study indicates that dentist's with the highest
income are usually located in central places in big cities. The findings
are in line with already exciting theories especially supporting the
gravitational models and agglomeration attributes. ?????? Proving a
central location within large cities are the best location to maximize
profit for Dental practices. ???

@@

Location theory gives regional economics its scientific disciplinary
identity and constitutes its theoretical methodological core Capello
(2011). It has typically microeconomic foundation and uses theoretical
models as well as adopting a statistical and geographical approach
Capello (2011). Furthermore, the theory uses the concept of
externalities in the spatial distribution of activities, thereby laying
the territorial bases for dynamic approach to economic growth Capello
(2011).

Regional growth theory involves spatial aspects of economic growth and
territorial distribution of income Capello (2011). It also involves
generating geographical advantages, in terms of easy or difficult access
to a particular area Capello (2011).

@@

Harald Hotelling's locational equilibrium is determined by a logic of
profit maximization whereby each producer controls its own market area.
Productivity advantages of cities and urban clusters with a high density
of firms increase profit by attracting a larger number of potential
customers, and more productive workers Capello (2011). Furthermore, the
attractiveness of a central location increase the cost of rent.

Alonso's bid rent model indicates the most profitable location for
firms. Closer to the centre with agglomeration attributes or in rural
areas with spacial monopoly and low rent. In gravitational models, the
attractiveness of the retail location, represent the size of the retail
centre. Furthermore, it depends on the variety of goods which can be
purchased at the same location McCann (2013).

The model of potential has the capacity to measure the potential of
attractiveness to a place. Bigger cities or more heavily populated areas
have a stronger gravitational force. A possible indicator to predict
places of growth potential for dentist practice or shopping malls
Capello (2015).

@@

The model assume that given the location of producers, and given demand
uniformly distributed geographically (in linear or circular form) the
market is divided into areas within each of which there operates a
single firm in a duopoly environment. Furthermore, no relocation costs
and demand only depend on location choice Capello (2015). The location
game starts off with the total market of AB, firm A in the middle of
location A and firm B starts in the middle of location B. One firm
starts relocating closer to the other to take some of the customers in
the other market area Capello (2015). The other respond by doing the
same and the game continues until both end up in the middle of the
market on the broader of AB. The end of the game is the position where
neither can increase sales volume by moving position Capello (2015).

@@

Marshall (1920) broadly divides externalities within agglomeration in
three main categories potentially drive sales. Firstly, knowledge
spill-over within industries or product specific technological
knowledge. Furthermore, market transactions in terms of value chain
transactions with industry-specialized buyers and suppliers. Lastly,
competition for specialized production factors such as labour and
product market competition Nielsen et al. (2021).

There are solidly established conclusions regarding the existence of
agglomeration economies Puga (2010). However, less proof of their
estimated magnitude. Hence, identifying the causes of agglomeration
economies, is proving more difficult Puga (2010). Nevertheless, there is
a large theoretical literature that develops these mechanisms Puga
(2010). Duranton and Puga (2004) discuss these classifications and
identify \emph{learning, sharing} and \emph{matching} as the main causes
of agglomeration economies.

A larger market allows for a more efficient sharing of local
infrastructure and facilities. Therefore, a variety of intermediate
input suppliers, or a pool of workers with similar skills Puga (2010).
Despite higher rent the dental industry and shopping malls, seams to
reap higher benefits in more populated areas as they are dependent on
being located where there is a higher volume of patients in order to
drive sales. The attraction for the consumers and users of public
facilities is overall cost reduction Puga (2010). Hence, the larger the
population sharing facilities the lower the cost per user Puga (2010).
Presumably, industrial factories and business clusters are more
dependent on being close to raw materials and industrial action.

Furthermore, a larger market also allows for a better matching between
employers and employees, improved chances of finding suitable and better
quality of matches Puga (2010). Shopping malls require skilled workers
to drive sales. However, they are not so dependent on highly educated
workers as dentists whom according to resent study, tend to prefer
bigger cities (as they are highly educated) Davis and Dingel (2020).
More-so, cites provide a constant market for specialized skills and more
productive work force Puga (2010). Perhaps, a possibility for higher
wages being compensated by more efficient workers in bigger
agglomeration economies. That said the Norwegian cities might differ
from the american cites in the study, as the population and clusters are
no where near the same size. Lastly, a larger market can also facilitate
learning by promoting the development and widespread adaptation of new
technologies and business practice Puga (2010).

Interactions with experienced workers helps acquire valuable skills.
Experienced workers remain in cities to share the rent of this learning
process Puga (2010). Besides this purposeful transition of knowledge,
the information literature on learning in cities has also emphasized the
unintended casual flow of information facilitated by big cities.'' Puga
(2010)

@@

The Alsonso model demonstrate geographical locations tied up to
location. Furthermore, It is an urbanized formulation of the nineteenth
century von Thünen model Capello (2015). In this case used to get an
indicating of where the most profitable location for dental practise.

Alonso assumes the existence of a city that cannot be build
instantaneously, and therefore of an effective rent-curve from the city
centre to the periphery. Furthermore, it determines the location for a
new firm willing to locate in the city and the profit the firm can
obtain. In some cases different from the normal or average price
competition Capello (2015).

As the Von Thünen assume a uniformed space where all land is equally
fertile, this model envisages a city, endowed with infrastructure which
cover the entire city in all directions whereas unit transport costs are
constant in space. More-so, the town or city has a single centre point.
The city centre of business district (CBD) is generally defined as the
most attractive for all firms. Capello (2015). The model also assume
perfect competition and unlimited demand, in other words it is
supply-orientated Capello (2015). Furthermore, demonstrating a specific
production function with fixed coefficients and constant returns to
scale.

Rent obtained in the model is the remainder left after transport,
production costs and profit has been subtracted from revenues Capello
(2015). The bid-rent curve will in this case demonstrate the prise of
rent within Bank and Financial locations, Retail locations, and
industrial locations.

@@

The illustration shows three categories: The most expensive rent closest
to the centre is the Bank and finance district, followed by the retail
agglomeration areas and industrial sites furthest away from the central
point. Dental practices are service orientated thus depending on
population density and central activities. Opening hours are relay
outside of office hours. Production cost does not appear to be important
in this case since dentist is a public service. Hence, not dependent of
regular transportation of sales goods. The rent realized by the
landowner is the envelope of the three rent supply curves Capello
(2015).

From a Marshallian way of thinking dentists should cluster somewhere in
this pink area to reap maximum sales volume. Furthermore, they may also
increase their profits further out in to other retail agglomeration
area. The attractiveness might be high in these places due to economies
of scope. As the positive benefits from the Marshallian agglomeration
attributes play out. Subsequently, increasing profit by trading higher
transportation cost for lower rent. In addition fewer dentists to
compete with supporting the hotellings game theory whereas companies
more to achieve spatial monopoly McCann (2013).

@@

The centre attracts firms, people and activities. Furthermore, it
influence the central points in diverse ways such as commuter movements,
diffusion of knowledge, corporations and network and personal
relationship Capello (2015). More-so, they are generated by the
intensity of the flows generated between places Capello (2015). The
gravitational pull to a place of interest for a dentist will determine
how much sales can be generated by locating there. Several locations can
then be measures against each other to determine profit maximisation.

These models has the potential to measures the future attractiveness of
a place. Changes in infrastructure such as easier access to a place
might change the attractiveness of a place or move the flow of people.
Subserviently, increasing the gravitationally pull between the two
places. Furthermore, The model uses population and distance to measure
the attractiveness and relations between two places Capello (2015).
Furthermore, these models have predictive capacity and can be used to
estimate the potential impact. In particular, location of new productive
activity in an area Capello (2015). For instance predict where to build
a shopping centre Capello (2015). By taking these factors in to account
dentist could predict the most profitable locations in years to come by
predicting the amount of people that will move from one city to another
or tow different points of interest within a city.

@@

Town and cities have long had shopping districts in which stores are
concentrated. To some extent such centre points were the result of
uncoordinated store-location decisions Brueckner (2011). Nevertheless,
not coincidentally according to Hotelling's theory (1929). He describes
the concentration of retail stores by a market desire to create a Nash
equilibrium. At this point all retailers expected sales maximisation
outcome cannot be improved by changing position. Contradictory to the
assumption that consumers will shop at the nearest retailer, they are
willing to travel significantly further to reach certain central places.
Marshall (1920) points out another explanation for this in his seminal
work on agglomeration. That being,intra-industry knowledge spillovers as
mechanisms leading to formation of industry cluster Nielsen et al.
(2021). In this case the dental industry information spillovers are
bounded in geographical space. Hence, the need to locate in close
proximity. That being the latest equipment, in order to work faster or
the right personnel to attract more patients. Another way to drive sales
in this area could be by dentist specialisation and patients referrals
between dentists.

Another important externalities expressed in Marshall (1920) theory is
market transactions. Whereas, geographical proximity reduces transaction
cost through several mechanism Nielsen et al. (2021). The market
transactions can be extended to the endogenous location of specialized
human capital Nielsen et al. (2021). ``It is well known among economists
and other social scientists that large cities have disproportionately
large shares of highly educated workers, and the trend has been growing
in recent decades.'' (\textbf{bakerBaker?}). Presumably, larger cities
would then be more attractive to dentists in terms of cost and quality
of specialised labour in smaller towns. Especially, those trying to
achieve spatial monopoly end up loosing profit by overpaying for labour
or in worst case scenario not finding enough dentists labour and end up
compromising on quality. In both cases growth potential would be
compromised.

Marshall also points out the competition within the geographical
location. Hence, the marketplace and business districts give consumers
low search and switching costs Capello (2015). The fact that owners of
stores seem to prefer highly concentrated shopping. More-so, it suggests
that the sales volumes outweigh the loss attributed to greater price
competition Capello (2015).

@@

Owners of shopping malls ``orchestrate'' the process of retail
agglomeration that happens naturally in towns and cities Brueckner
(2011). Naturally, the same retail agglomeration and Nash equilibrium
may occur in shopping malls if not more intensely as the shops are in
closer proximity. Furthermore, the potential for more knowledge
spillovers, higher volume of market transaction and more competition.
When retail agglomeration is orchestrated by the owner of a shopping
mall, the strength and direction of such externalities are taken into
account Brueckner (2011). Much so, because the price of rent is
dependent on the overall revenue of a shopping mall. Hence, the owner of
the mall wants to choose the mix of stores and their sizes taking these
externalities, attributes and anchor stores attractiveness into account
Brueckner (2011).

@@

Spatial data analysis is a rapidly growing area in Statistics Haslett
(1992). The Journal of the statistics society define such data as being
objects in space, which objects have physical location. Furthermore, the
analyses are spatial if these locations are relevant to the
interpretation of the data Haining, Wise, and Ma (1998).

The examination of the data involves the examination of data collected
from q-gis maps;
\texttt{it\ will\ involve\ little\ more.\ Graphical\ techniques\ for\ examining\ such\ maps\ are\ thus\ a\ central\ part\ of\ the\ methodology.\ A\ number\ of\ such\ techniques\ are\ discussed."}
Haslett (1992).

@@

The bid rent models are under the assumption that Dentists use spatial
completion to maximise profit. Nevertheless, the cost of dental care in
Norway can vary depending on a number of other factors such as the type
of treatment, the location of the dental practice and the dentist
experience and qualifications. The Norwegian Dental Association rate the
average cost of check ups is between four to eight hundred Norwegian
Krone's (\textbf{tannlege?}). Bertrand model also take price as
competitive advantage in to consideration.

@@

Shopping malls have a relatively short history in Norway, with the first
shopping mall opening in the late sixties. Prior to the most retail
activities took place in the city centre. Furthermore, these malls
typically featured a mix of shops, restaurants and entertainment
options, and they were often located in suburban areas or on the
outskirts of cities (\textbf{home-n?}). In resent years, there has been
a trend towards larger and more comprehensive shopping malls in Norway
(\textbf{home-n?}).

However, the growth of shopping malls in Norway has not been without
controversy. Some critics argue that the development of large shopping
malls has led to the decline of traditional shopping streets and local
businesses, while other raise concerns about the environmental impact of
these large scale commercial developments (\textbf{kanvin2014?}).

Information about Wine monopoly locations has also been studied in this
paper. As the wine monopoly are anchor stores they may affect the
attractiveness of the shopping mall (\textbf{kanvin2014?}). The
Norwegian government regulates the placements of wine monopoly stores in
shopping malls to prevent completion with town centres
(\textbf{omsorgsdepartementet2022?}). More importantly, the government
may take in to account the potential impact of exciting retail centres
in the region. This regional economic approach aim to promote
sustainable development and balance economic growth across the country
(\textbf{kanvin2014?}).

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-brueckner2011}{}}%
Brueckner, Jan K. 2011. \emph{Lectures on Urban Economics}. MIT Press.

\leavevmode\vadjust pre{\hypertarget{ref-capello2011}{}}%
Capello, Roberta. 2011. {``{[}PDF{]}location, Regional Growth and Local
Development Theories,''} January.
\url{https://www.academia.edu/48547951/_PDF_Location_Regional_Growth_and_Local_Development_Theories}.

\leavevmode\vadjust pre{\hypertarget{ref-capello2015}{}}%
---------. 2015. \emph{Regional Economics}. 2nd ed. Routledge.
\url{https://doi.org/10.4324/9781315720074}.

\leavevmode\vadjust pre{\hypertarget{ref-davis2020}{}}%
Davis, Donald R., and Jonathan I. Dingel. 2020. {``The Comparative
Advantage of Cities.''} \emph{Journal of International Economics} 123
(March): 103291. \url{https://doi.org/10.1016/j.jinteco.2020.103291}.

\leavevmode\vadjust pre{\hypertarget{ref-duranton2004}{}}%
Duranton, Gilles, and Diego Puga. 2004. {``Chapter 48 -
Micro-Foundations of Urban Agglomeration Economies.''} In \emph{Handbook
of Regional and Urban Economics}, edited by J. Vernon Henderson and
Jacques-François Thisse, 4:2063--2117. Cities and Geography. Elsevier.
\url{https://doi.org/10.1016/S1574-0080(04)80005-1}.

\leavevmode\vadjust pre{\hypertarget{ref-haining1998}{}}%
Haining, R., S. Wise, and J. Ma. 1998. {``Exploratory Spatial Data
Analysis.''} \emph{Journal of the Royal Statistical Society: Series D
(The Statistician)} 47 (3): 457--69.
\url{https://doi.org/10.1111/1467-9884.00147}.

\leavevmode\vadjust pre{\hypertarget{ref-haslett1992}{}}%
Haslett, John. 1992. {``Spatial Data Analysis---Challenges.''}
\emph{Journal of the Royal Statistical Society: Series D (The
Statistician)} 41 (3): 271--84. \url{https://doi.org/10.2307/2348549}.

\leavevmode\vadjust pre{\hypertarget{ref-mccann2013}{}}%
McCann, Philip. 2013. \emph{Modern Urban and Regional Economics}. Second
edition. Oxford University Press.

\leavevmode\vadjust pre{\hypertarget{ref-nielsen2021}{}}%
Nielsen, Bo Bernhard, Christian Geisler Asmussen, Cecilie Dohlmann
Weatherall, and Ditte Håkonsson Lyngemark. 2021. {``Marshall Vs Jacobs
Agglomeration and the Micro-Location of Foreign and Domestic Firms.''}
\emph{Cities} 117 (October): 103322.
\url{https://doi.org/10.1016/j.cities.2021.103322}.

\leavevmode\vadjust pre{\hypertarget{ref-puga2010}{}}%
Puga, Diego. 2010. {``THE MAGNITUDE AND CAUSES OF AGGLOMERATION
ECONOMIES.''} \emph{Journal of Regional Science} 50 (1): 203--19.
\url{https://doi.org/10.1111/j.1467-9787.2009.00657.x}.
